%! Author = trdng
%! Date = 3/18/2026

% Preamble
\documentclass[11pt]{article}

% Packages
\usepackage{amsmath}
\usepackage{braket}
\usepackage{indentfirst}

% Document
\begin{document}

    \section*{Single Qubit Gates}
    \indent Single qubit gates are unitary operations that act on one qubit at a time.
    Mathematically, they are $(2 \times 2)$ unitary matrices that transforms a qubit state vector.
    \newline\indent
    A general qubit state is $\ket{\psi} = \alpha\ket{0} + \beta\ket{1} $, with $|\alpha|^2 + |\beta|^2 = 1$.
    \newline\indent
    A single qubit gate $U$ acts as $ \ket{\psi'} = U \ket{\psi}$

    \subsection{Hadamard}\label{subsec:hadamard}

    Hadamard converts phase info into measurable bit flips.
    It acts like a perfect quantum coin toss.
    A matrix represents Hadamard gate as follow

    \[
    H = \frac{1}{\sqrt{2}}
    \begin{pmatrix}
        1 & 1 \\
        1 & -1
    \end{pmatrix}
    \]

    When applying on a qubit, it results

    \[
        H\ket{0} = \frac{\ket{0} + \ket{1}}{\sqrt{2}} and
        H\ket{1} = \frac{\ket{0} - \ket{1}}{\sqrt{2}}
    \]

    \subsection{X-gate}\label{subsec:x-gate}

    X-gate flips the qubit.
    A matrix represents X-gate as follow

    \[
        X =
        \begin{pmatrix}
            0 & 1 \\
            1 & 0
        \end{pmatrix}
    \]

    When applying on a qubit, it results

    \[
        X\ket{0} = \ket{1} and
        X\ket{1} = \ket{0}
    \]

    In Bloch sphere, it is the rotation of 180° around the x-axis.

    \subsection{Y-gate}\label{subsec:y-gate}

    A matrix represents Y-gate is as follows

    \[
        Y = \begin{pmatrix}
                0 & -i \\
                i & 0
        \end{pmatrix}
    \]

    When applying on a qubit, it results

    \[
        Y\ket{0} = i\ket{1} and
        Y\ket{1} = -i\ket{0}
    \]

    In Bloch sphere, it is the rotation of 180° around the y-axis.

    \subsection{Z-gate}\label{subsec:z-gate}

    A matrix represents Y-gate is as follows

    \[
        Y = \begin{pmatrix}
                1 & 0 \\
                0 & -1
        \end{pmatrix}
    \]

    When applying on a qubit, it results

    \[
        Z\ket{0} = \ket{0} and
        Z\ket{1} = -\ket{1}
    \]

    In Bloch sphere, it is the rotation of 180° around the z-axis.

    \subsection{Phase-gate}\label{subsec:phase-gate}

    \[
        Y = \begin{pmatrix}
                1 & 0 \\
                0 & i
        \end{pmatrix}
    \]

    When applying on a qubit, it results

    \[
        S\ket{0} = \ket{0} and
        S\ket{1} = i\ket{1}
    \]

    Phase-gate adds 90° phase to $\ket{1}$

\end{document}